Online Laterality Battery Pre-Processing - short version

### 2020-02-02 battery-preprocessing_v3_DB.Rmd
Found some errors in the finger task analysis so corrected these, and also added an analysis of SDs of RTs on that task. Clarified that for chimeric task, accuracy is not % correct, because both L and R correct, so this is a choice measure. But for RDT, we do have percent correct, and this is now renamed as p.corr to make that more explicit.

    temp_dat$subject[myrow] <- subname # add row for subject
    temp_dat$side[myrow] <- latlist[j] # add row for side
    temp_dat$accurate[myrow] <- 100*mean(tmp1$Correct,na.rm=TRUE) # add row for accurate responses - this has been modified to give % correct
    temp_dat$RT[myrow] <- mean(tmp2$RT,na.rm=TRUE) # add row for mean RT
    temp_dat$N[myrow]<-length(tmp1$RT) # add count of rows for this subject
  }
}
return(temp_dat)
}
```

@@

#Now repeat for day 2
origdat <- chim_dat2
origdat$X1 <- chim_dat2$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Chimeric faces day 2'

dopirate(origdat,task,measure)
origdat$X1 <- chim_dat2$RT
measure <- 'RT'
dopirate(origdat,task,measure)

@@

w<-which(RDT_df1$accurate<60)
if (length(w)>0)
{RDT_df1$accurate[w]<-NA
RDT_df1$RT[w]<-NA
}
w<-which(RDT_df2$accurate<60)
if (length(w)>0)
{RDT_df2$accurate[w]<-NA
RDT_df2$RT[w]<-NA
  }
```

@@

#now create LI
origdat$X1 <- origdat$accurate
mycolname<-"RDT_acc_LI"
polarity<- 1 #so LI is positive in predicted direction
mycutoff <- 75
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"RDT_RT_LI"
polarity <- (-1)
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

#Cutoff has removed those with low accuracy, but we should also remove their RT LIs
#We do this by finding those with NA in accuracy LI
w<-which(is.na(demographics$RDT_acc_LI))
if(length(w)>0)
{demographics$RDT_RT_LI[w]<-NA}
#t.test to test if laterality index differs from zero
t.test(demographics$RDT_acc_LI)
t.test(demographics$RDT_RT_LI)
#rename column to identify day
myncol <- length(colnames(demographics))
colnames(demographics)[(myncol-1):myncol] <- paste0('Day',d,'.',colnames(demographics)[(myncol-1):myncol])

}
cor1<-cor(demographics$Day1.RDT_acc_LI,demographics$Day2.RDT_acc_LI,use='complete.obs')
cor2<-cor(demographics$Day1.RDT_RT_LI,demographics$Day2.RDT_RT_LI,use='complete.obs')
plot(demographics$Day1.RDT_acc_LI,demographics$Day2.RDT_acc_LI,main='Rhyme detect: LI for accuracy')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor1,3)))
plot(demographics$Day1.RDT_RT_LI,demographics$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor2,3)))
```
### T-test as alternative to LI for RDT

@@

```{r readdataFinger, warning=FALSE}
for (d in 1:2){ #we will process day 1 and day 2 together
  origfile <- all_dat1
  if (d==2)
  {origfile <- all_dat2}
#As with earlier reading in, we default to saving finger2 (day2), but at end of the chunk we reassign to finger1 in the first run through the d loop
task <- "Key Sequence Task"
wanted <- c("subject", "Response", "display") # select wanted columns
finger2 <- procdata(origfile,wanted,task)

# count subjects
nsub <-length(levels(finger2$subject))
print(paste('day',d,task,": subjects: ", nsub))

if (d==1)
{finger1 <- finger2
 }
}
```

@@

```{r scorefinger}
# We can go directly to creation of a data.frame by subjects, and use string matching to identify N correct seqs
# y, m, i, b- complex
# w, v, r, x- complex
# t, u, m, b- easy
# w, r, v, x- easy

@@

finger.df <- data.frame(matrix(NA,nrow=nsub,ncol=9))
colnames(finger.df)<-c('subject','Leasy1','Reasy1','Lhard1','Rhard1',
                       'Leasy2','Reasy2','Lhard2','Rhard2')
for (n in 1:nsub){
  finger.df$subject[n] <- subids[n]
  myc <- 1 #will be incremented so sums written to correct columns
  for (d in 1:2){ #each day data
    myfinger <- finger1
    if(d==2){
      myfinger<- finger2}
      tempfinger <- filter(myfinger,subject==subids[n],Response %in% c('y','m','i','b','t','u','w','v','r','x')) #exclude rows with other stuff in Response
      #(this is not essential as nonsequences will just get ignored)
      wholeseq <-paste0(tempfinger$Response,collapse='') #make a long string of responses
      for (s in 1:4){
        myc<-myc+1 #increment column counter
        target <- myseq[s]
        finger.df[n,myc]<-str_count(wholeseq,target) #count how many target sequences are in the long string that contains all responses
      }
    }
  }
#Need same data in long form for pirate plot, so reorganise
finger.left <- finger.df[,c(1,2,4,6,8)]
finger.right <- finger.df[,c(1,3,5,7,9)]
finger.left$side<-'Left'
finger.right$side <-'Right'
colnames(finger.right)<-c('subject','easy1','hard1','easy2','hard2','side')
colnames(finger.left)<-colnames(finger.right)
finger.long <-rbind(finger.left,finger.right)

@@

measure <- 'hard2'
origdat$X1 <- origdat$hard2 #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Finger hard day 2'
dopirate(origdat,task,measure)
```

@@

cor1 <- cor(demographics$Day1_finger.easy_LI,demographics$Day2_finger.easy_LI)
plot(demographics$Day1_finger.easy_LI,demographics$Day2_finger.easy_LI,main='Finger easy LI')
abline(v=0)
abline(h=0)
text(40,10,paste0('r = ',round(cor1,3)))
cor2 <- cor(demographics$Day1_finger.hard_LI,demographics$Day2_finger.hard_LI)
plot(demographics$Day1_finger.hard_LI,demographics$Day2_finger.hard_LI,main='Finger hard LI')
abline(v=0)
abline(h=0)
text(40,10,paste0('r = ',round(cor2,3)))

>>>>

    temp_dat$subject[myrow] <- subname # add row for subject
    temp_dat$side[myrow] <- latlist[j] # add row for side
    temp_dat$accurate[myrow] <- sum(tmp$Correct,na.rm=TRUE) # add row for N accurate response
    temp_dat$p.corr[myrow] <- 100*mean(tmp$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice. For RDT, can make error of selecting wrong side, so % correct is meaningful

    temp_dat$RT[myrow] <- mean(tmp2$RT,na.rm=TRUE) # add row for mean RT
    temp_dat$N[myrow]<-length(tmp1$RT) # add count of rows for this subject
  }
}
return(temp_dat)
}
```

@@

#Now repeat for day 2
origdat <- chim_dat2
origdat$X1 <- chim_dat2$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Chimeric faces day 2'
measure <- 'Accuracy'
dopirate(origdat,task,measure)
origdat$X1 <- chim_dat2$RT
measure <- 'RT'
dopirate(origdat,task,measure)

@@

w<-which(RDT_df1$p.corr<60)
if (length(w)>0)
{RDT_df1$accurate[w]<-NA
RDT_df1$RT[w]<-NA
}
w<-which(RDT_df2$p.corr<60)
if (length(w)>0)
{RDT_df2$accurate[w]<-NA
RDT_df2$RT[w]<-NA
  }
```

@@

#now create LI
origdat$X1 <- origdat$p.corr
mycolname<-"RDT_p.corr_LI"
polarity<- 1 #so LI is positive in predicted direction
mycutoff <- 75
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"RDT_RT_LI"
polarity <- (-1)
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

#Cutoff has removed those with low accuracy, but we should also remove their RT LIs
#We do this by finding those with NA in accuracy LI
w<-which(is.na(demographics$RDT_acc_LI))
if(length(w)>0)
{demographics$RDT_RT_LI[w]<-NA}
#t.test to test if laterality index differs from zero
t.test(demographics$RDT_p.corr_LI)
t.test(demographics$RDT_RT_LI)
#rename column to identify day
myncol <- length(colnames(demographics))
colnames(demographics)[(myncol-1):myncol] <- paste0('Day',d,'.',colnames(demographics)[(myncol-1):myncol])

}
cor1<-cor(demographics$Day1.RDT_p.corr_LI,demographics$Day2.RDT_p.corr_LI,use='complete.obs')
cor2<-cor(demographics$Day1.RDT_RT_LI,demographics$Day2.RDT_RT_LI,use='complete.obs')
plot(demographics$Day1.RDT_p.corr_LI,demographics$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor1,3)))
plot(demographics$Day1.RDT_RT_LI,demographics$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor2,3)))
```
### T-test as alternative to LI for RDT

@@

```{r readdataFinger, warning=FALSE}
for (d in 1:2){ #we will process day 1 and day 2 together
  origfile <- all_dat1
  if (d==2)
  {origfile <- all_dat2}
#As with earlier reading in, we default to saving finger2 (day2), but at end of the chunk we reassign to finger1 in the first run through the d loop
task <- "Key Sequence Task"
wanted <- c("subject", "Response", "display","RT") # select wanted columns
#NB RT is not meaningful, in that we are just counting N sequences in fixed time.
# But it could be looked at for measure of variability.
finger2 <- procdata(origfile,wanted,task)
finger2$complex <-'easy'
task <- "Key Sequence Task (wider keys)"
finger2a<-procdata(origfile,wanted,task)
finger2a$complex <-'hard'
finger2<-rbind(finger2,finger2a)
# count subjects
nsub <-length(levels(finger2$subject))
print(paste('day',d,task,": subjects: ", nsub))

finger2$RTbase[2:nrow(finger2)]<-finger2$RT[1:(nrow(finger2)-1)]
finger2$RTa <- finger2$RT-finger2$RTbase
#remove negative or v long RTs - can indicate when block changes etc
w<-c(which(finger2$RTa<0),which(finger2$RTa>1000))
finger2$RTa[w]<-NA
if (d==1)
{finger1 <- finger2
 }
}
```

@@

Added exploratory analysis of SDs of RTs for keypress rather than means, but this was not particularly illuminating.

```{r scorefinger}
# We can go directly to creation of a data.frame by subjects, and use string matching to identify N correct seqs
# y, m, i, b- complex R
# t, u, m, b- easy R
# w, v, r, x- complex L
# w, r, v, x- easy L

@@

finger.df <- data.frame(matrix(NA,nrow=nsub,ncol=17))
colnames(finger.df)<-c('subject','Rhard1','Reasy1','Lhard1','Leasy1',
                       'Rhard2','Reasy2','Lhard2','Leasy2',
                       'Rhard1sd','Reasy1sd','Lhard1sd','Leasy1sd',
                       'Rhard2sd','Reasy2sd','Lhard2sd','Leasy2sd')
for (n in 1:nsub){
  finger.df$subject[n] <- subids[n]
  myc <- 1 #will be incremented so sums written to correct columns
  for (d in 1:2){ #each day data
    myfinger <- finger1
    if(d==2){
      myfinger<- finger2}
      tempfinger <- filter(myfinger,subject==subids[n],Response %in% c('y','m','i','b','t','u','w','v','r','x')) #exclude rows with other stuff in Response
      #(this is not essential as nonsequences will just get ignored)

      wholeseq <-paste0(tempfinger$Response,collapse='') #make a long string of responses
      for (s in 1:4){
        myc<-myc+1 #increment column counter
        target <- myseq[s]
        finger.df[n,myc]<-str_count(wholeseq,target) #count how many target sequences are in the long string that contains all responses

      }
      righteasy<-filter(tempfinger,display=='right hand',complex=='easy')
      righthard<-filter(tempfinger,display=='right hand',complex=='hard')
      lefteasy<-filter(tempfinger,display=='left hand',complex=='easy')
      lefthard<-filter(tempfinger,display=='left hand',complex=='hard')
      startcol<-10
      if(d==2){startcol<-14}
      finger.df[n,startcol]<-sd(righthard$RTa,na.rm=T)
      finger.df[n,(startcol+1)]<-sd(righteasy$RTa,na.rm=T)
      finger.df[n,(startcol+2)]<-sd(lefthard$RTa,na.rm=T)
      finger.df[n,(startcol+3)]<-sd(lefteasy$RTa,na.rm=T)
    }
}
# Substitute NA for any totals less than 5 - suggest using wrong keys perhaps

for (c in 2:9){
  w<-which(finger.df[,c]<5)
  if (length(w)>0){
     finger.df[w,2:17]<-NA

}
#Need same data in long form for pirate plot, so reorganise
finger.left <- finger.df[,c(1,4,5,8,9,12,13,16,17)]
finger.right <- finger.df[,c(1,2,3,6,7,10,11,14,15)]
finger.left$side<-'Left'
finger.right$side <-'Right'
colnames(finger.right)<-c('subject','hard1','easy1','hard2','easy2',
                          'hard1sd','easy1sd','hard2sd','easy2sd','side')
colnames(finger.left)<-colnames(finger.right)
finger.long <-rbind(finger.left,finger.right)

@@

measure <- 'hard2'
origdat$X1 <- origdat$hard2 #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Finger hard day 2'
dopirate(origdat,task,measure)

measure<-'easy1sd'
origdat$X1 <- origdat$easy1sd #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
task <- 'Finger easy day 1sd'
dopirate(origdat,task,measure)

measure <- 'easy2sd'
origdat$X1 <- origdat$easy2sd #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Finger easy day 2'
dopirate(origdat,task,measure)

measure <- 'hard1sd'
origdat$X1 <- origdat$hard1sd #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Finger hard day 1'
dopirate(origdat,task,measure)

measure <- 'hard2sd'
origdat$X1 <- origdat$hard2sd #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Finger hard day 2'
dopirate(origdat,task,measure)
```

@@

cor1 <- cor(demographics$Day1_finger.easy_LI,demographics$Day2_finger.easy_LI,use='complete.obs')
plot(demographics$Day1_finger.easy_LI,demographics$Day2_finger.easy_LI,main='Finger easy LI')
abline(v=0)
abline(h=0)
text(40,10,paste0('r = ',round(cor1,3)))
cor2 <- cor(demographics$Day1_finger.hard_LI,demographics$Day2_finger.hard_LI,use='complete.obs')
plot(demographics$Day1_finger.hard_LI,demographics$Day2_finger.hard_LI,main='Finger hard LI')
abline(v=0)
abline(h=0)
text(40,10,paste0('r = ',round(cor2,3)))
